### 2023-04-11 article.tex
Merge single-line paragraphs

To preserve code for the longest time, 
both code and its dependencies can be put into a so-called container.

The most reproducible way of submitting the code of an experiment,
is to put all code with all the (software) dependencies
in a file that acts as a virtual computational environment,
called a 'virtual container' (from now on: 'container').
Such a container is close to the golden standard of reproducible research
as suggested by 
\cite{peng2011reproducible} 
\iffalse
(see also Figure \ref{fig:peng2011reproducible})
\fi
.

@@

Useful paradata, in general, are (1) comprehensive,
(2) documented appropriately, (3) documented in co-evolution with the data,
and (4) friendly to computers (\cite{huvila2022improving}).
In this section, these ideal properties of paradata are applied to code.

However, code has multiple uses,
as code can be used to (1) reproduce, (2) replicate, or (3)
extend a computational experiment (\cite{benureau2018re}).
Depending on the intended use of the code,
there are different requirements for code being useful paradata.

@@

For code to be ideal paradata, it must be usefully 
documented (\cite{huvila2022improving}).

For purposes of reproducing code,
it should at least be documented how to run the code and what it ought to do.
Although this may be obvious, only 57\% out of 56 Science papers 
with obtainable code (in total there were 180 papers) 
do so (\cite{stodden2018empirical}).

@@

Code should be distributed in standard ways (\cite{peng2006reproducible}),
as is done by using a code hosting website
(see subsection \ref{subsec:code-hosting}).

% \paragraph{Code must have extensive error handling}

Code must be more extensive when it is (intended to be) 
used on different data,
as then 'code must act as a teacher for future developers' (\cite{sadowski2018modern}).
Error handling is one of the mechanisms to do so.
In genetic epidemiology, it is common to have incomplete or missing data,
so analyses should take this into account with clear error messages.

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code with all its (software) dependencies 
in a container.

% \paragraph{Runnable code can precisely reproduce an experiment}

Containers allow a computation experiment to be highly reproducible:
given the same data, an experiment put into a container will give
the same results on different platforms, at least in theory.
In practice, differences may be observed when peripheral factors
are different, such as the random numbers as generated by an operating
system, or data that are downloaded from online (and hence, probably changing) sources.

@@

A first example is to use a tool to enforce a coding style 
(e.g. the Tidyverse style guide (\cite{wickham2019advanced}) for R,
or PEP 8 (\cite{van2001pep}) for Python),
as following a consistent coding style improves software quality (\cite{fang2001}).

% \paragraph{Using a tool for complexity}

A second example is to use a tool to enforce a low cyclomatic complexity.
The cyclomatic complexity is approximately defined 
as the number of independent paths that
the code can be executed. 
\iffalse
For example, code that has only one \verb|if| statement
has a cyclomatic complexity of 2, as the condition within the \verb|if|
statement can be true or false,
resulting in the body of the \verb|if| statement being either
executed or ignored.
\fi
The cyclomatic complexity correlates with code complexity,
where more complex code is likelier to contain or give rise to bugs 
(\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}).

@@

Before applying these features, the first step is to publish 
the code.

When applying these general recommendations to code, 
this list can be phrased more precisely:

>>>>

To preserve code for the longest time, 
both code and its dependencies can be put into a so-called container.
The most reproducible way of submitting the code of an experiment,
is to put all code with all the (software) dependencies
in a file that acts as a virtual computational environment,
called a 'virtual container' (from now on: 'container').
Such a container is close to the golden standard of reproducible research
as suggested by 
\cite{peng2011reproducible} 
\iffalse
(see also Figure \ref{fig:peng2011reproducible})
\fi
.

@@

Useful paradata, in general, are (1) comprehensive,
(2) documented appropriately, (3) documented in co-evolution with the data,
and (4) friendly to computers (\cite{huvila2022improving}).
In this section, these ideal properties of paradata are applied to code.
However, code has multiple uses,
as code can be used to (1) reproduce, (2) replicate, or (3)
extend a computational experiment (\cite{benureau2018re}).
Depending on the intended use of the code,
there are different requirements for code being useful paradata.

@@

For code to be ideal paradata, it must be usefully 
documented (\cite{huvila2022improving}).
For purposes of reproducing code,
it should at least be documented how to run the code and what it ought to do.
Although this may be obvious, only 57\% out of 56 Science papers 
with obtainable code (in total there were 180 papers) 
do so (\cite{stodden2018empirical}).

@@

Code should be distributed in standard ways (\cite{peng2006reproducible}),
as is done by using a code hosting website
(see subsection \ref{subsec:code-hosting}).
% \paragraph{Code must have extensive error handling}
Additionally, code must be more extensive when it is (intended to be) 
used on different data,
as then 'code must act as a teacher for future developers' (\cite{sadowski2018modern}).
Error handling is one of the mechanisms to do so.
In genetic epidemiology, it is common to have incomplete or missing data,
so analyses should take this into account with clear error messages.

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code with all its (software) dependencies 
in a container.
% \paragraph{Runnable code can precisely reproduce an experiment}
Containers allow a computation experiment to be highly reproducible:
given the same data, an experiment put into a container will give
the same results on different platforms, at least in theory.
In practice, differences may be observed when peripheral factors
are different, such as the random numbers as generated by an operating
system, or data that are downloaded from online (and hence, probably changing) sources.

@@

A first example is to use a tool to enforce a coding style 
(e.g. the Tidyverse style guide (\cite{wickham2019advanced}) for R,
or PEP 8 (\cite{van2001pep}) for Python),
as following a consistent coding style improves software quality (\cite{fang2001}).
% \paragraph{Using a tool for complexity}
A second example is to use a tool to enforce a low cyclomatic complexity.
The cyclomatic complexity is approximately defined 
as the number of independent paths that
the code can be executed. 
\iffalse
For example, code that has only one \verb|if| statement
has a cyclomatic complexity of 2, as the condition within the \verb|if|
statement can be true or false,
resulting in the body of the \verb|if| statement being either
executed or ignored.
\fi
The cyclomatic complexity correlates with code complexity,
where more complex code is likelier to contain or give rise to bugs 
(\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}).

@@

Before applying these features, the first step is to publish 
the code. 
When applying these general recommendations to code, 
this list can be phrased more precisely:
